---
title: "Objects preparation"
author: "Rasmus Rydbirk"
date: "2024-11-11"
output: 
  html_document:
    toc: True
    toc_float: True
---

# Setup

```{r setup, message=FALSE}
library(pagoda2)
library(conos)
library(qs)
library(dplyr)
library(magrittr)
library(scHelper) # github.com/rrydbirk/scHelper
```

# Conos objects

## Load annotation

```{r}
anno.major <- qread("anno_major.qs")
```

## con_major

The easiest way to extract the list with raw count matrices is to extract it from the Conos object that we provide for all cells. We do this first and then preprocess.

```{r}
con.major <- qread("con_major.qs", nthreads = 10)

cms.preprocessed <- con.major$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(basicP2proc, make.geneknn = F, get.tsne = F, get.largevis = F, n.cores = 32)
```

Then, we can re-create the Conos object. Please note, we provide additional embeddings in the prepared object that you'll lose if you save this object. Also, the embedding may not be exactly the same.

After the next step, we use the major object to create objects for subpopulations.

```{r, eval = F}
con.major <- cms.preprocessed %>%  
  Conos$new(n.cores = 32)

con.major$buildGraph()
con.major$findCommunities()
con.major$embedGraph(method = "UMAP", embedding.name = "UMAP_refined7", min.dist = 0.0001, n.neighbors = 120, spread = 5)
con.major$embedGraph(method = "largeVis", embedding.name = "largeVis") # Not necessary to run, we used the UMAP embedding

qsave("con_major.qs", nthreads = 10)
```

## con_adipocytes

```{r, eval = F}
con <- con.major$samples %>% 
  lapply('[[', "misc") %>% 
  lapply('[[', "rawCounts") %>% 
  lapply(Matrix::t) %>% 
  lapply(\(cm) cm[, colnames(cm) %in% names(anno.major)[anno.major == "Adipocytes"]]) %>% 
  lapply(basicP2proc, make.geneknn = F, get.tsne = F, get.largevis = F, n.cores = 32) %>% 
  Conos$new(n.cores = 32)

con$buildGraph()
con$findCommunities()
con$embedGraph(method = "UMAP", embedding.name = "UMAP")
con$embedGraph(method = "largeVis", embedding.name = "largeVis")
con$findCommunities(resolution = 0.5, name = "leiden0.5")
con$findCommunities(resolution = 1.5, name = "leiden1.5")

qsave("con_adipocytes.qs", nthreads = 10)
```

## con_adipocytes_aspc_macrophages

```{r, eval = F}
anno.immune <- qread("anno_immune.qs") %>% 
  .[. %in% c("ATM", "Early LAM", "LAM", "mono/mac")]

anno.aspc <- qread("anno_aspc.qs")

anno <- qread("anno_major_final.qs") %>% 
  .[!names(.) %in% names(anno.immune)] %>% 
  .[!names(.) %in% names(anno.aspc)] %>% 
  {factor(c(., anno.immune, anno.aspc))}

cts <- c("ASPC_CXCL14", "ASPC_PPARG", "ASPC_EPHA3", "ASPC_DPP4", "Adipocytes", "ATM","Early LAM", "LAM", "mono/mac")

anno <- anno[anno %in% cts] %>% 
  factor()

qsave(anno, "anno_adipocytes_aspc_macrophages.qs")

con <- con.major$samples %>% 
  lapply('[[', "misc") %>% 
  lapply('[[', "rawCounts") %>% 
  lapply(Matrix::t) %>% 
  lapply(\(cm) cm[, colnames(cm) %in% names(anno)[anno %in% cts]])

con %<>% lapply(basicP2proc, make.geneknn = F, get.tsne = F, get.largevis = F, n.cores = 32) %>% 
  Conos$new(n.cores = 32)

con$buildGraph()
con$findCommunities()
con$embedGraph(method = "UMAP", embedding.name = "UMAP")
con$embedGraph(method = "largeVis", embedding.name = "largeVis")

qsave(con, "con_adipocytes_aspc_macrophages.qs")
```

## con_aspc

```{r, eval = F}
con <- con.major$samples %>% 
  lapply('[[', "misc") %>% 
  lapply('[[', "rawCounts") %>% 
  lapply(Matrix::t) %>% 
  lapply(\(cm) cm[, colnames(cm) %in% names(anno.major)[anno.major == "ASPC"]]) %>% 
  lapply(basicP2proc, make.geneknn = F, get.tsne = F, get.largevis = F, n.cores = 32) %>% 
  Conos$new(n.cores = 32)

con$buildGraph()
con$findCommunities()
con$embedGraph(method = "UMAP", embedding.name = "UMAP")
con$embedGraph(method = "largeVis", embedding.name = "largeVis")
con$findCommunities(resolution = 0.5, name = "leiden0.5")
con$findCommunities(resolution = 1.5, name = "leiden1.5")

qsave(con, "con_aspc.qs", nthreads = 10)
```

## con_myeloid

```{r, eval = F}
con <- con.major$samples %>% 
  lapply('[[', "misc") %>% 
  lapply('[[', "rawCounts") %>% 
  lapply(Matrix::t) %>% 
  lapply(\(cm) cm[, colnames(cm) %in% names(anno.major)[anno.major == "Myeloid immune cells")]]) %>% 
  lapply(basicP2proc, make.geneknn = F, get.tsne = F, get.largevis = F, n.cores = 32) %>% 
  Conos$new(n.cores = 32)

con$buildGraph()
con$findCommunities()
con$embedGraph(method = "UMAP", embedding.name = "UMAP")
con$embedGraph(method = "largeVis", embedding.name = "largeVis")
con$findCommunities(resolution = 0.5, name = "leiden0.5")
con$findCommunities(resolution = 1.5, name = "leiden1.5")

qsave(con, "con_myeloid.qs", nthreads = 10)
```

## con_lymphoid

```{r, eval = F}
con <- con.major$samples %>% 
  lapply('[[', "misc") %>% 
  lapply('[[', "rawCounts") %>% 
  lapply(Matrix::t) %>% 
  lapply(\(cm) cm[, colnames(cm) %in% names(anno.major)[anno.major == "Lymphoid immune cells"]]) %>% 
  .[sapply(., ncol) > 30] %>% 
  lapply(basicP2proc, make.geneknn = F, get.tsne = F, get.largevis = F, n.cores = 32, nPcs = 20) %>% 
  Conos$new(n.cores = 32)

con$buildGraph()
con$findCommunities()
con$embedGraph(method = "UMAP", embedding.name = "UMAP")
con$embedGraph(method = "largeVis", embedding.name = "largeVis")
con$findCommunities(resolution = 0.5, name = "leiden0.5")
con$findCommunities(resolution = 1.5, name = "leiden1.5")

qsave(con, "con_lymphoid.qs", nthreads = 10)
```

## con_immune

```{r, eval = F}
con <- con.major$samples %>% 
  lapply('[[', "misc") %>% 
  lapply('[[', "rawCounts") %>% 
  lapply(Matrix::t) %>% 
  lapply(\(cm) cm[, colnames(cm) %in% names(anno.major)[anno.major %in% c("Lymphoid immune cells", "Myeloid immune cells")]]) %>% 
  lapply(basicP2proc, make.geneknn = F, get.tsne = F, get.largevis = F, n.cores = 32) %>% 
  Conos$new(n.cores = 32)

con$buildGraph()
con$findCommunities()
con$embedGraph(method = "UMAP", embedding.name = "UMAP")
con$embedGraph(method = "largeVis", embedding.name = "largeVis")
con$findCommunities(resolution = 0.5, name = "leiden0.5")
con$findCommunities(resolution = 1.5, name = "leiden1.5")

qsave(con, "con_immune.qs", nthreads = 10)
```

## con_vascular

```{r, eval = F}
con <- con.major$samples %>% 
  lapply('[[', "misc") %>% 
  lapply('[[', "rawCounts") %>% 
  lapply(Matrix::t) %>% 
  lapply(\(cm) cm[, colnames(cm) %in% names(anno.major)[anno.major %in% c("ECs", "SMCs", "Lymphatic ECs", "Pericytes")]]) %>% 
  lapply(basicP2proc, make.geneknn = F, get.tsne = F, get.largevis = F, n.cores = 32, nPcs = 50) %>% 
  Conos$new(n.cores = 32)

con$buildGraph()
con$findCommunities()
con$embedGraph(method = "UMAP", embedding.name = "UMAP")
con$embedGraph(method = "largeVis", embedding.name = "largeVis")
con$findCommunities(resolution = 0.5, name = "leiden0.5")
con$findCommunities(resolution = 1.5, name = "leiden1.5")

qsave(con, "con_vascular.qs", nthreads = 10)
```

# Export data to Python

## aspc_adi_macro objects

```{r}
con <- qread("con_adipocytes_aspc_macrophages.qs", nthreads = 10)
annotation <- qread("anno_adipocytes_aspc_macrophages.qs")

prepareObjectsForPython(con, 
                        annotation,
                        "",
                        "aspc_adi_macro",
                        "largeVis")
```

## major objects

```{r}
con <- qread("con_major.qs", nthreads = 10)
annotation <- qread("anno_major.qs")

prepareObjectsForPython(con, 
                        annotation,
                        "",
                        "major",
                        "UMAP")

# Visits
spc <- con$getDatasetPerCell()

spc %>%
  {data.frame(
    visit = unname(.) %>% 
      as.character() %>% 
      grepl.replace(patterns = c("vis1","vis2","vis3"), c("visit1","visit2","visit3")), 
    cellids = names(.))} %>% 
  write.table("all_cells.visits", 
              sep = ",", 
              dec = ".", 
              row.names = F, 
              col.names = F)
```
